Comparison of approaches to using machine learning models to increase the efficiency of the global search algorithm for solving multicriterial problems

### 2024-06-30 2024/OPTIMA/OPTIMA_2024_MCO.tex
before submission

After the construction of the separating hyperplane, for each trial point the $d_i$ distance is calculated and normalized according to the following expression:
\begin{equation}
    \label{eq:17}
d'_i=
\begin{cases}
  \begin{matrix}
     d_i / d_{max}, & d_i > 0 \\
     -d_i / d_{min}, & d_i < 0 
  \end{matrix}
\end{cases}, 
0 \leq i \leq k-1,
\end{equation}
where $d_i$ is the distance from the $i$-th point to the hyperplane, $d_{max}$, $d_{min}$ are maximum and minimum distances to the hyperplane.

The value $R_{ps}(i)$ from (\ref{eq:16}) for the interval $(x_{i-1},x_{i})$ is calculated as the sum of normalized distances (\ref{eq:17}) to points $x_i$ and $x_{i-1}$. Let's denote this version of GSA as ML\_MGSA\_Dist.

@@

As part of the proposed approach, the machine learning model is built according to the rules similar to the construction of the SVC model from section \ref{ssec:41}. As before, all trial points are divided into two classes -- points of the Pareto set and points that do not belong to the Pareto set. The classifier is trained on the selected classes. Note that the classifier models of most machine learning libraries can return not only the class label, but also the probability of relevance to the class. Unlike the approach described in Subsection \ref{ssec:41} the characteristic $R_{ps}(i)$ is computed as the average of probabilities of situations when the points $x_i$ and $x_{i-1}$ from interval $(x_{i-1},x_{i})$ belong to the Pareto set.

@@

Second, the probabilities undergo the transformation
\begin{equation}
    \label{eq:18}
p'_i= \frac{ \log (p_i) - \log (p_{min})}{ \log (p_{max}) - \log (p_{min})} , 0 \leq i \leq k-1,
\end{equation}
where $p_i$ is a probability of the event that the point $x_i$ belongs to the Pareto set, $p_{min}$, $p_{max}$ are minimum and maximum values of such probabilities. After transformation the function $R_{ps}(i)$ is calculated, as earlier, as the average value of probabilities of events in which the points $x_i$ and $x_{i-1}$ from interval $(x_{i-1}, x_i)$ belong to the Pareto set.

@@

Computations were conducted on the supercomputer ``Lobachevsky'' of the Nizhni Novgorod State University (operating system CentOS 7, control system SLURM). To obtain the executable program code the compilers Intel C++ 17.0.0 and python 3.9 were used. Computations were performed with the support of the package scikit-learn 0.24.2 and the optimization software iOpt \cite{iOpt_url}.

>>>>

After the construction of the separating hyperplane, for each trial point the $d_i$ distance is calculated and normalized according to the following expression:
\begin{equation}
    \label{eq:17}
d'_i=
\begin{cases}
  \begin{matrix}
     d_i / d_{max}, & d_i > 0 \\
     -d_i / d_{min}, & d_i < 0 
  \end{matrix}
\end{cases}, 
0 \leq i \leq k-1,
\end{equation}
where $d_i$ is the distance from the $i$-th point to the hyperplane, $d_{max}$, $d_{min}$ are maximum and minimum distances to the hyperplane.
The value $R_{ps}(i)$ from (\ref{eq:16}) for the interval $(x_{i-1},x_{i})$ is calculated as the sum of normalized distances (\ref{eq:17}) to $i$-th and $(i-1)$-th points. Let's denote this version of GSA as ML\_MGSA\_Dist.

@@

As part of the proposed approach, the machine learning model is built according to the rules similar to the construction of the SVC model from section \ref{ssec:41}. As before, all trial points are divided into two classes -- points of the Pareto set and points that do not belong to the Pareto set. The classifier is trained on the selected classes. Note that the classifier models of most machine learning libraries can return not only the class label, but also the probability of relevance to the class. Unlike the approach described in Subsection \ref{ssec:41} the characteristic $R_{ps}(i)$ is computed as the average of probabilities of situations when the points
%$y_i=y(x_i)$ and $y_{i-1}=y(x_{i-1})$
correspond to boudaries of the interval $(x_{i-1},x_{i})$ belong to the Pareto set.

@@

Second, the probabilities undergo the transformation
\begin{equation}
    \label{eq:18}
p'_i= \frac{ \log (p_i) - \log (p_{min})}{ \log (p_{max}) - \log (p_{min})} , 0 \leq i \leq k-1,
\end{equation}
where $p_i$ is a probability of the event that the point $x_i$ belongs to the Pareto set, $p_{min}$, $p_{max}$ are minimum and maximum values of such probabilities. After transformation the function $R_{ps}(i)$ is calculated, as earlier, as the average value of probabilities of events in which the points correspond to bounaries of the interval $(x_{i-1}, x_i)$ belong to the Pareto set.

@@

Computations were conducted on the ``Lobachevsky'' supercomputer node (two Intel Sandy Bridge E5-2660 2.2 GHz processors and 64 Gb RAM) using the scikit-learn 0.24.2 library.
The proposed algorithms were implemented in Python in intelligent optimization framework iOpt \cite{iOpt_url}.

%Computations were performed with the support of the scikit-learn 0.24.2 framework.

%of the Nizhni Novgorod State University (operating system CentOS 7, control system SLURM).To obtain the executable program code the compilers Intel C++ 17.0.0 and python 3.9 were used.  and the optimization software iOpt \cite{iOpt_url}.
